Data Import, Clean, and Process

### 2025-03-11 notebooks/01-data-import-clean-process.qmd
Removed and added a slew of files

---
title: Data Import, Clean, and Process
author: Isaiah
format: 
  html: 
    code-fold: false
    code-tools: true
    code-overflow: wrap
    code-copy: true
    code-annotations: true
---

@@

@@

# load raw SPSS export (includes display order vars)
# downloaded on 2024-10-29 at 12:02 PM
raw_spss <- haven::read_sav("data-raw/VMF_Election_Worker_Recruitment_Survey_with_display_order_20241029_T12.02.sav")

@@

A data dictionary contains metadata about the data. The `labelled::generate_dictionary` function is used to create a data dictionary. A dataframe is loaded into the R environment with the number of observations equal to number of variables in the original data set. One problem that I encountered was that I was unable to quickly save the data dictionary generated using the `labelled` R package. As an alternative, I utilized the `data_dict()` function from the R package `surveytoolbox` in order to save the dictionary as a `.csv` file.

```{r}
#| label: raw-data-dict

# data dictionary from labelled package. Won't save as .csv file
raw_spss_dict <- raw_spss |> labelled::generate_dictionary()

# data dictionary from labelled package. Will save as .csv file
raw_spss_dict <- raw_spss |> surveytoolbox::data_dict()

@@

The data dictionary is the same as a codebook, but I also wanted to have a codebook that contained only the survey questions, variable labels (i.e., question text), responses (value labels), and associated values. So the saved `codebook` contains only the survey questions and the two grouping variables (`qset` and `group`).

>>>>

---
title: Data Import, Clean, and Process
format: 
  bookup-html:
    code-fold: false
    code-tools: true
    code-overflow: wrap
    code-copy: true
    code-annotations: true
---

## Set up

@@

```{r}
#| label: setup
#| include: true
#| eval: true
#| echo: true

set.seed(1234)

# load packages
library(tidyverse)

@@

# load raw SPSS export (includes display order vars)
# downloaded on 2024-10-29 at 12:02 PM
raw_spss <- haven::read_sav("data-raw/vmf_election_worker_recruitment_survey_raw_spss_sanitized_20241029_T12.02.sav")

@@

A data dictionary contains metadata about the data. The `labelled::generate_dictionary` function is used to create a data dictionary (alternatively, use the `data_dict()` function from R package [surveytoolbox](https://martinctc.github.io/surveytoolbox/)). This function produces a dataframe into the R environment with the number of observations (rows) equal to number of variables (columns) in the original data set. For ease, I prefer to save the codebook/data dictionary as a `.csv` file. In order to do so, ensure that the function `labelled::convert_list_columns_to_character()` is run on the data dictionary which allows you to save the data dictionary as a `.csv` file. Alternatively, the
function `surveytoolbox::data_dict()` permits saving as a .csv file without the
additional step.

Also note that I use the native R pipe format ` |> ` (tip: try hotkey `CTRL + .`), which is essentially equivalent to the `magrittr` forward-pipe operator ` %>% `.

```{r}
#| label: raw-data-dict
#| echo: true

# data dictionary from labelled package.
raw_spss_dict <- raw_spss |> 
  labelled::generate_dictionary() |> 
  # this allows me to save dictionary as .csv file
  labelled::convert_list_columns_to_character()

# alternative
# raw_spss_dict <- raw_spss |> surveytoolbox::data_dict()

@@

The data dictionary is the same as a codebook, but I also wanted to have a codebook that contained **only** the survey questions, variable labels (i.e., question text), responses (value labels), and associated values. So the saved `codebook` contains only the survey questions and the two grouping variables (`qset` and `group`).
